---
title: "Primary Analysis for the Daily Death Rate"
author: "JLee"
date: "5/21/2020"
output: html_document
---

```{r,message=FALSE,include=FALSE,echo=FALSE}
library(mgcv);library(dlnm);library(gamm4);library(lme4);library(splines);library(Hmisc);
library(lubridate);library(dplyr);library(stringr);library(RCurl);library(httr);
library(dplyr);library(MASS);

library(boot)
library(table1)

# Data import
mydata <-  read.csv("Data/us_state_daily_death_JHU.csv")[,-1]
mydata$date <- as.Date(mydata$date)
state_predic <- read.csv("Data/us_state_predictors.csv")[,-1]
state_tpredic <-  read.csv("Data/us_state_timevarying_daily_predictors.csv")[,-1]
state_tpredic$date <- as.Date(state_tpredic$date)
policy <-  read.csv("Data/state_daily_Shome_policy.csv")[,-1]
policy$date <- as.Date(policy$date)
policy <- policy %>% group_by(state) %>% mutate(shome_since_first = cumsum(stayathome))
policy <- as.data.frame(policy) 

mydata <- merge(mydata, state_predic, by=c("state"))
mydata$date <- as.Date(mydata$date)
mydata <- merge(mydata, policy, by=c("state", "date"))
mydata <- mydata[mydata$cum_death>0, ]
mydata$daily_death <- ifelse(mydata$daily_death<0, 0, mydata$daily_death)
mydata <- mydata[order(mydata$state, as.Date(mydata$date)),]

# create average for temperature and humidity

mydata$temp = (mydata$summer_tmmx + mydata$winter_tmmx)/2
mydata$humid = (mydata$mean_summer_rm + mydata$mean_winter_rm)/2

```


```{r, include=FALSE, echo=FALSE}
# TABLE 1
#demographic_state[, 3:23] = sapply(demographic_state[, 3:23], as.numeric)

table1(~   Age_gp2 + Age_gp3 + Age_gp4 + 
         + density
         + temp + humid 
         + Bachelor_orhigher_perc 
         + poverty_perc
         + Black.perc + Hispanic.or.Latino.perc
         + bed_rate + smoke_perc + obese_perc +inactive_perc + diabetes_perc + uninsured_perc, data=state_predic)

```

```{r, echo=FALSE}


## Final data
#state_tpredic$date <- as.Date(state_tpredic$date)
#mydata <- merge(mydata, state_tpredic, by=c("state", "date"))
#mydata$date <- as.Date(mydata$date)

# create lag for relative testing rate
lagk <- 15
mydata$rtest <- mydata$test/(mydata$population/1000)

for (kk in (1:lagk)){
  create = paste0("test_days_ago",kk)
  for(i in (1:length(unique(mydata$state)))){
    dates_before = mydata[which(mydata$state == unique(mydata$state)[i]),]$date-kk
    ind = match(dates_before,mydata[which(mydata$state == unique(mydata$state)[i]),]$date)
    mydata[which(mydata$state==unique(mydata$state)[i]),create]=mydata[which(mydata$state==unique(mydata$state)[i]),][ind,]$rtest
  }
}


# create lag for positive testing rate

lagk <- 15
state_tpredic$test_3ave <- frollmean(state_tpredic[, "test"], 3)
state_tpredic$pos_test_3ave <- frollmean(state_tpredic[, "pos_test"], 3)

state_tpredic$rptest_3ave <- state_tpredic$pos_test_3ave/(state_tpredic$test_3ave)

for (kk in (1:lagk)){
  create = paste0("ptest_3ave_days_ago",kk)
  for(i in (1:length(unique(state_tpredic$state)))){
    dates_before = state_tpredic[which(state_tpredic$state == unique(state_tpredic$state)[i]),]$date-kk
    ind = match(dates_before,state_tpredic[which(state_tpredic$state == unique(state_tpredic$state)[i]),]$date)
    state_tpredic[which(state_tpredic$state==unique(state_tpredic$state)[i]),create]=state_tpredic[which(state_tpredic$state==unique(state_tpredic$state)[i]),][ind,]$rptest_3ave
  }
}

state_tpredic$date <- as.Date(state_tpredic$date)
mydata <- merge(mydata, state_tpredic, by=c("state", "date"))
mydata$date <- as.Date(mydata$date)

lagk <- 15
mydata$test_3ave <- frollmean(mydata[, "test"], 3)
mydata$pos_test_3ave <- frollmean(mydata[, "pos_test"], 3)

mydata$rptest_3ave <- mydata$pos_test_3ave/(mydata$test_3ave)

for (kk in (1:lagk)){
  create = paste0("ptest_3ave_days_ago",kk)
  for(i in (1:length(unique(mydata$state)))){
    dates_before = mydata[which(mydata$state == unique(mydata$state)[i]),]$date-kk
    ind = match(dates_before,mydata[which(mydata$state == unique(mydata$state)[i]),]$date)
    mydata[which(mydata$state==unique(mydata$state)[i]),create]=mydata[which(mydata$state==unique(mydata$state)[i]),][ind,]$rptest_3ave
  }
}


#mydata$rtest <- mydata$rtest/(mydata$population/1000)
#mydata$test_weeks_ago1 <- mydata$test_weeks_ago1/(mydata$population/1000)
#mydata$test_weeks_ago2 <- mydata$test_weeks_ago2/(mydata$population/1000)

# create PCA for the movement data

mydata2 <- na.omit(mydata)
mydata2 <- mydata2[mydata2$state!="Maine",]  ## Maine update the negative test results weekly or biweekly

# create # of days since first death (in our dataset)

mydata2$days_since_first = 1 
for (i in 1:length(unique(mydata2$state))){
  for (k in 2:nrow(mydata[which(mydata2$state==unique(mydata2$state)[i]),])){
  mydata2[which(mydata2$state==unique(mydata2$state)[i]),"days_since_first"][k]=
    (mydata2[which(mydata2$state==unique(mydata2$state)[i]),"date"][k]-mydata2[which(mydata2$state==unique(mydata2$state)[i]),"date"][1])
  }
}

tmp <- mydata2[is.na(mydata2$parks), ]
tmp$date <- as.Date(tmp$date)
for(i in 1:nrow(tmp)){
 mydata2[mydata2$state==tmp$state[i] & mydata2$date==tmp$date[i], "parks"] <-  (mydata2[mydata2$state==tmp$state[i] & mydata2$date==tmp$date[i]-1, "parks"] + mydata2[mydata2$state==tmp$state[i] & mydata2$date==tmp$date[i]+1, "parks"])/2
}

pca_move <- prcomp(mydata2[,c("retail_rec", "grocery_pharmacy", "parks", "transit", "work", "residential")], center=TRUE, scale.=TRUE)
mydata2 <- cbind(mydata2, pca_move$x)



mydata3 <- mydata2[mydata2$rptest>=0, ]


#mydata3 <- mydata2[mydata2$rptest!=1, ] ## Remove the first week of Ohio and Maryland due to they have 0 negative test. 

```

# Model 1

(1) Daily_death ~ s(days_since_first) + scaled independent variables
```{r,echo=FALSE}

mod1 = gamm4(daily_death~ s(days_since_first,k=8,bs='cr') + factor(q_popdensity)
             + scale(Age_gp1) + scale(Age_gp3) + scale(Age_gp4) 
             + scale(poverty_perc) #+ scale(log(susceptible_perc))
             + scale(Black.perc) + scale(Hispanic.or.Latino.perc)
             + scale(temp) + scale(humid)
             + scale(Bachelor_orhigher_perc) 
             + scale(bed_rate) + scale(smoke_perc) + scale(obese_perc) + scale(diabetes_perc)
             + offset(log(population)),
             random = ~(1|state), family=negbin(1), data = mydata3)

summary(mod1$gam)
plot(mod1$gam, trans=exp, ylab="Mortality Rate Ratios", xlab="Days since the first death incidence")
```

(2) Daily_death ~ s(days_since_first) + scaled independent variables + relative_test with 15 days lag

```{r,echo=FALSE}

mod2 = gamm4(daily_death~ s(days_since_first,k=8,bs='cr') + factor(q_popdensity)
             + scale(Age_gp1) + scale(Age_gp3) + scale(Age_gp4) 
             + scale(poverty_perc) 
             + scale(Black.perc) + scale(Hispanic.or.Latino.perc)
             + scale(temp) + scale(humid)
             + scale(Bachelor_orhigher_perc) 
             + scale(bed_rate) + scale(smoke_perc) + scale(obese_perc)  + scale(diabetes_perc)
             + scale(rtest) + scale(test_days_ago1) + scale(test_days_ago2) + scale(test_days_ago3) + scale(test_days_ago4)
             + scale(test_days_ago5) + scale(test_days_ago6) + scale(test_days_ago7) + scale(test_days_ago8) + scale(test_days_ago9)
             + scale(test_days_ago10) + scale(test_days_ago11) + scale(test_days_ago12) + scale(test_days_ago13) + scale(test_days_ago14) + scale(test_days_ago15)
             + offset(log(population)),
             random = ~(1|state), family=negbin(1), data = mydata3)

summary(mod2$gam)
plot(mod2$gam, trans=exp, ylab="Mortality Rate Ratios", xlab="Days since the first death incidence", select=1)
```



(3) Daily_death ~ s(days_since_first) + scaled independent variables + positive_test_rate with 15 days lag

```{r,echo=FALSE}

mod3 = gamm4(daily_death~ s(days_since_first,k=8,bs='cr') + factor(q_popdensity)
             + scale(Age_gp1) + scale(Age_gp3) + scale(Age_gp4) 
             + scale(poverty_perc) 
             + scale(Black.perc) + scale(Hispanic.or.Latino.perc)
             + scale(temp) + scale(humid)
             + scale(Bachelor_orhigher_perc) 
             + scale(bed_rate) + scale(smoke_perc) + scale(obese_perc)  + scale(diabetes_perc)
             + scale(rptest) + scale(ptest_days_ago1) + scale(ptest_days_ago2) + scale(ptest_days_ago3) + scale(ptest_days_ago4)
             + scale(ptest_days_ago5) + scale(ptest_days_ago6) + scale(ptest_days_ago7) + scale(ptest_days_ago8) + scale(ptest_days_ago9)
             + scale(ptest_days_ago10) + scale(ptest_days_ago11) + scale(ptest_days_ago12) + scale(ptest_days_ago13) + scale(ptest_days_ago14) + scale(ptest_days_ago15)
             + offset(log(population)),
             random = ~(1|state), family=negbin(1), data = mydata3)

summary(mod3$gam)
plot(mod3$gam, trans=exp, ylab="Mortality Rate Ratios", xlab="Days since the first death incidence", select=1)
plot(0:15, exp(mod3$gam$coefficients[grep("ptest", names(mod3$gam$coefficients))]), type="l", xlab="Lag", ylab="Mortality Rate Ratios")
```

(4) Daily_death ~ s(days_since_first) + scaled independent variables + positive_test_rate with 15 days lag

```{r,echo=FALSE}

## Only include counties having more than lagk days observation period

Testing <- as.matrix(mydata3[, c(lages)])
lagpoint <- matrix(rep(seq(0, lagk), nrow(mydata3)), nrow=nrow(mydata3), byrow=T)

kk <- 8
mod4 = gamm4(daily_death~ s(days_since_first,k=8,bs='cr') + factor(q_popdensity)
             + scale(Age_gp1) + scale(Age_gp3) + scale(Age_gp4) 
             + scale(poverty_perc) 
             + scale(Black.perc) + scale(Hispanic.or.Latino.perc)
             + scale(temp) + scale(humid)
             + scale(Bachelor_orhigher_perc) 
             + scale(bed_rate) + scale(smoke_perc) + scale(obese_perc)  + scale(diabetes_perc)
             + s(lagpoint, by=Testing, bs="cr", k=kk) +
             + offset(log(population)),
             random = ~(1|state), family=negbin(1), data = mydata3)

summary(mod4$gam)
plot(mod4$gam, trans=exp, ylab="Mortality Rate Ratios", xlab="Days since the first death incidence", select=1)
plot(mod4$gam, trans=exp, ylab="Mortality Rate Ratios", xlab="Lages", select=2)
```



(5) Daily_death ~ s(days_since_first) + scaled independent variables + relative_test with 15 days lag + movement

```{r,echo=FALSE}

mod4 = gamm4(daily_death~ s(days_since_first,k=8,bs='cr') + factor(q_popdensity)
             + scale(Age_gp1) + scale(Age_gp3) + scale(Age_gp4) 
             + scale(poverty_perc) 
             + scale(Black.perc) + scale(Hispanic.or.Latino.perc)
             + scale(temp) + scale(humid)
             + scale(Bachelor_orhigher_perc) 
             + scale(bed_rate) + scale(smoke_perc) + scale(obese_perc)  + scale(diabetes_perc)
             + scale(rptest) + scale(ptest_days_ago1) + scale(ptest_days_ago2) + scale(ptest_days_ago3) + scale(ptest_days_ago4)
             + scale(ptest_days_ago5) + scale(ptest_days_ago6) + scale(ptest_days_ago7) + scale(ptest_days_ago8) + scale(ptest_days_ago9)
             + scale(ptest_days_ago10) + scale(ptest_days_ago11) + scale(ptest_days_ago12) + scale(ptest_days_ago13) + scale(ptest_days_ago14) + scale(ptest_days_ago15)
             + work 
             + offset(log(population)),
             random = ~(1|state), family=negbin(1), data = mydata3)

summary(mod4$gam)
plot(mod4$gam, trans=exp, ylab="Mortality Rate Ratios", xlab="Days since the first death incidence", select=1)
```

```{r, echo=FALE}
mod6 = gamm4(daily_death~ s(days_since_first,k=8,bs='cr') + factor(q_popdensity)
             + scale(Age_gp1) + scale(Age_gp3) + scale(Age_gp4) 
             + scale(poverty_perc) 
             + scale(Black.perc) + scale(Hispanic.or.Latino.perc)
             + scale(temp) + scale(humid)
             + scale(Bachelor_orhigher_perc) 
             + scale(bed_rate) + scale(smoke_perc) + scale(obese_perc)  + scale(diabetes_perc)
             + rptest
             + offset(log(population)),
             random = ~(1|state), family=negbin(1), data = mydata3)

summary(mod6$gam)
plot(mod6$gam, trans=exp, ylab="Mortality Rate Ratios", xlab="Days since the first death incidence", select=1)
```

```{r, echo=FALE}
mod7 = gamm4(daily_death~ s(days_since_first,k=8,bs='cr') + factor(q_popdensity)
             + scale(Age_gp1) + scale(Age_gp3) + scale(Age_gp4) 
             + scale(poverty_perc) 
             + scale(Black.perc) + scale(Hispanic.or.Latino.perc)
             + scale(temp) + scale(humid)
             + scale(Bachelor_orhigher_perc) 
             + scale(bed_rate) + scale(smoke_perc) + scale(obese_perc)  + scale(diabetes_perc)
             + rptest_3ave + scale(ptest_3ave_days_ago1) + scale(ptest_3ave_days_ago2) + scale(ptest_3ave_days_ago3) + scale(ptest_3ave_days_ago4)
             + scale(ptest_3ave_days_ago5) + scale(ptest_3ave_days_ago6) + scale(ptest_3ave_days_ago7) + scale(ptest_3ave_days_ago8) + scale(ptest_3ave_days_ago9)
             + scale(ptest_3ave_days_ago10) + scale(ptest_3ave_days_ago11) + scale(ptest_3ave_days_ago12) + scale(ptest_3ave_days_ago13) + scale(ptest_3ave_days_ago14) + scale(ptest_3ave_days_ago15)

             + offset(log(population)),
             random = ~(1|state), family=negbin(1), data = mydata2)

summary(mod7$gam)
plot(mod7$gam, trans=exp, ylab="Mortality Rate Ratios", xlab="Days since the first death incidence", select=1)

lages <- c("rptest_3ave", paste("ptest_3ave_days_ago", seq(1:lagk), sep=""))
Testing <- as.matrix(mydata3[, c(lages)])
lagpoint <- matrix(rep(seq(0, lagk), nrow(mydata3)), nrow=nrow(mydata3), byrow=T)

kk <- 8

mod8 = gamm4(daily_death~ s(days_since_first,k=8,bs='cr') + factor(q_popdensity)
             + scale(Age_gp1) + scale(Age_gp3) + scale(Age_gp4) 
             + scale(poverty_perc) 
             + scale(Black.perc) + scale(Hispanic.or.Latino.perc)
             + scale(temp) + scale(humid)
             + scale(Bachelor_orhigher_perc) 
             + scale(bed_rate) + scale(smoke_perc) + scale(obese_perc)  + scale(diabetes_perc)
             + s(lagpoint, by=Testing, bs="cr", k=10) +
             + offset(log(population)),
             random = ~(1|state), family=negbin(1), data = mydata3)

summary(mod8$gam)
plot(mod8$gam, trans=exp, ylab="Mortality Rate Ratios", xlab="Days since the first death incidence", select=1)
plot(mod8$gam, ylab="Mortality Rate Ratios", xlab="Lages", select=2)
```
```{r, echo=FALSE}

mydata3$sipo <- 0
mydata3$sipo[mydata3$shome_since_first>1 & mydata3$shome_since_first<=7] <- 1
mydata3$sipo[mydata3$shome_since_first>7 & mydata3$shome_since_first<=14] <- 2
mydata3$sipo[mydata3$shome_since_first>14 & mydata3$shome_since_first<=21] <- 3
mydata3$sipo[mydata3$shome_since_first>21 & mydata3$shome_since_first<=28] <- 4
mydata3$sipo[mydata3$shome_since_first>28 & mydata3$shome_since_first<=35] <- 5
mydata3$sipo[mydata3$shome_since_first>35] <- 6

mydata3$Bachelor_orhigher_perc <- as.numeric(mydata3$Bachelor_orhigher_perc)

mod7 =gamm4(daily_death~ s(days_since_first, k=8, bs="cr")  #factor(q_popdensity)
             + scale(Age_gp1) + scale(Age_gp3) + scale(Age_gp4) 
             + scale(poverty_perc) 
             + scale(Black.perc) + scale(Hispanic.or.Latino.perc)
            # + scale(temp) + scale(humid)
             + scale(Bachelor_orhigher_perc) 
            # + scale(bed_rate) #+ scale(smoke_perc) + scale(obese_perc)  + scale(diabetes_perc)
             + factor(sipo)
             + offset(log(population)),
             random=~(1|state), family=negbin(1), data = mydata3)
summary(mod7$gam)
plot(mod7$gam, trans=exp, ylab="Mortality Rate Ratios", xlab="Days since the first death incidence", select=1)


mod7 = gamm4(daily_death~ s(days_since_first,k=8,bs='cr') + factor(q_popdensity)
             + scale(Age_gp1) + scale(Age_gp3) + scale(Age_gp4) 
             + scale(poverty_perc) 
             + scale(Black.perc) + scale(Hispanic.or.Latino.perc)
             + scale(temp) + scale(humid)
             + scale(Bachelor_orhigher_perc) 
             + scale(bed_rate) + scale(smoke_perc) + scale(obese_perc)  + scale(diabetes_perc)
             + rptest_3ave + scale(ptest_3ave_days_ago1) + scale(ptest_3ave_days_ago2) + scale(ptest_3ave_days_ago3) + scale(ptest_3ave_days_ago4)
             + scale(ptest_3ave_days_ago5) + scale(ptest_3ave_days_ago6) + scale(ptest_3ave_days_ago7) + scale(ptest_3ave_days_ago8) + scale(ptest_3ave_days_ago9)
             + scale(ptest_3ave_days_ago10) + scale(ptest_3ave_days_ago11) + scale(ptest_3ave_days_ago12) + scale(ptest_3ave_days_ago13) + scale(ptest_3ave_days_ago14) + scale(ptest_3ave_days_ago15) + factor(sipo)
             + offset(log(population)),
             random = ~(1|state), family=negbin(1), data = mydata3)


```
