---
title: "Primary_Analysis_one_lag"
author: "JLee"
date: "5/23/2020"
output:
  pdf_document: default
  html_document: default
---
```{r, include=FALSE, echo=FALSE}
library(mgcv);library(gamm4);library(ggplot2)
library(lubridate);library(dplyr);
library(boot)
library(table1)

# Data import
mydata <-  read.csv("Data/us_state_weekly_death_JHU.csv")[,-1]
mydata$date <- as.Date(mydata$date)
state_predic <-  read.csv("Data/us_state_predictors.csv")[,-1]
#state_tpredic <-  read.csv("Data/us_state_timevarying_weekly_predictors.csv")[,-1]
#state_tpredic$date <- as.Date(state_tpredic$date)

movement <- read.csv("Data/us_state_weekly_movement.csv")[,-1]
testing <- read.csv("Data/weekly_testing.csv")[,-1]
policy <-  read.csv("Data/state_weekly_policy.csv")[,-1]
movement$date <- as.Date(movement$date)
testing$date <- as.Date(testing$date)
policy$date <- as.Date(policy$date)

mydata <- merge(mydata, state_predic, by=c("state"))
mydata$date <- as.Date(mydata$date)
mydata <- mydata[order(mydata$state, as.Date(mydata$date)),]


# create lag for positive testing rate
lagk <- 4
testing$rptest <- testing$pos_test/(testing$test)*100

for (kk in (1:lagk)){
  create = paste0("ptest_weeks_ago",kk)
  for(i in (1:length(unique(testing$state)))){
    dates_before = testing[which(testing$state == unique(testing$state)[i]),]$date-7*kk
    ind = match(dates_before,testing[which(testing$state == unique(testing$state)[i]),]$date)
    testing[which(testing$state==unique(testing$state)[i]),create]=testing[which(testing$state==unique(testing$state)[i]),][ind,]$rptest
  }
}

# create lag for resident
lagk <- 4

for (kk in (1:lagk)){
  create = paste0("resident_weeks_ago",kk)
  for(i in (1:length(unique(movement$state)))){
    dates_before = movement[which(movement$state == unique(movement$state)[i]),]$date-7*kk
    ind = match(dates_before,movement[which(movement$state == unique(movement$state)[i]),]$date)
    movement[which(movement$state==unique(movement$state)[i]),create]=movement[which(movement$state==unique(movement$state)[i]),][ind,]$residential
  }
}

## Merging with time-varying variables - testing, movement, policy
mydata2 <- merge(mydata, testing, by=c("state", "date"))
mydata2 <- merge(mydata2, movement, by=c("state", "date"))

policy2 <- policy[, !names(policy) %in% c("Start.Date", "End.Date")]
mydata2 <- merge(mydata2, policy2, by=c("state", "date"))

# create # of weeks since stay-at-home policy (in our dataset)

mydata2 <- mydata2 %>% group_by(state) %>% mutate(shome_since_first = cumsum(stayathome))
mydata2 <- as.data.frame(mydata2)

mydata2$rtest <- mydata2$test/(mydata2$population/1000)
#mydata$test_weeks_ago1 <- mydata$test_weeks_ago1/(mydata$population/1000)
#mydata$test_weeks_ago2 <- mydata$test_weeks_ago2/(mydata$population/1000)

# create PCA for the movement data
mydata2 <- mydata2[, !names(mydata) %in% c("Start.Date", "End.Date")]
mydata2 <- na.omit(mydata2)
mydata2 <- mydata2[mydata2$state!="Maine",]  ## Maine update the negative test results weekly or biweekly
# create # of weeks since first death (in our dataset)

mydata2$weeks_since_first = 1 
for (i in 1:length(unique(mydata2$state))){
  for (k in 2:nrow(mydata2[which(mydata2$state==unique(mydata2$state)[i]),])){
  mydata2[which(mydata2$state==unique(mydata2$state)[i]),"weeks_since_first"][k]=
    (mydata2[which(mydata2$state==unique(mydata2$state)[i]),"date"][k]-mydata2[which(mydata2$state==unique(mydata2$state)[i]),"date"][1])/7+1
  }
}

pca_move <- prcomp(mydata2[,c("retail_rec", "grocery_pharmacy", "parks", "transit", "work", "residential")] # make sure col# correct
                   , center=TRUE, scale.=TRUE)
pca_move
summary(pca_move)
mydata3 <- cbind(mydata2, pca_move$x)
```

# Model 2
(1) Weekly_death ~  s(days_since_first) + scaled independent variables + the current positive test ratio + PC1 + PC2 + Mask indicator + random intercept

```{r,echo=FALSE}

mod1 = gamm4(weekly_death~ s(weeks_since_first, k=8, bs="cr") + factor(q_popdensity)
             + scale(Age_gp1) + scale(Age_gp3) + scale(Age_gp4) 
             + scale(poverty_perc) 
             + scale(Black.perc) + scale(Hispanic.or.Latino.perc)
             + scale(temp) + scale(perc)
             + scale(Bachelor_orhigher_perc) 
             + scale(bed_rate)  + scale(smoke_perc) + scale(obese_perc)  
             + rptest 
             + PC1 + PC2
             + factor(Mask)
             + offset(log(population)),
             random = ~(1|state), family=negbin(1), data = mydata3)

summary(mod1$gam)
plot(mod1$gam, trans=exp, ylab="Mortality Rate Ratios", xlab="Weeks since the first death incidence", select=1)
```

(2) Weekly_death ~  s(days_since_first) + scaled independent variables + the 1 week lag positive test ratio + PC1 + PC2 + Mask indicator + random intercept

```{r,echo=FALSE}

mod2 = gamm4(weekly_death~ s(weeks_since_first, k=8, bs="cr") + factor(q_popdensity)
             + scale(Age_gp1) + scale(Age_gp3) + scale(Age_gp4) 
             + scale(poverty_perc) 
             + scale(Black.perc) + scale(Hispanic.or.Latino.perc)
             + scale(temp) + scale(perc)
             + scale(Bachelor_orhigher_perc) 
             + scale(bed_rate)  + scale(smoke_perc) + scale(obese_perc)  
             + ptest_weeks_ago1
             + PC1 + PC2
             + factor(Mask)
             + offset(log(population)),
             random = ~(1|state), family=negbin(1), data = mydata3)

summary(mod2$gam)
plot(mod2$gam, trans=exp, ylab="Mortality Rate Ratios", xlab="Weeks since the first death incidence", select=1)
```

(3) Weekly_death ~  s(days_since_first) + scaled independent variables + the 2 week lag positive test ratio + PC1 + PC2 + Mask indicator + random intercept

```{r,echo=FALSE}

mod3 = gamm4(weekly_death~ s(weeks_since_first, k=8, bs="cr") + factor(q_popdensity)
             + scale(Age_gp1) + scale(Age_gp3) + scale(Age_gp4) 
             + scale(poverty_perc) 
             + scale(Black.perc) + scale(Hispanic.or.Latino.perc)
             + scale(temp) + scale(perc)
             + scale(Bachelor_orhigher_perc) 
             + scale(bed_rate)  + scale(smoke_perc) + scale(obese_perc)  
             + ptest_weeks_ago2
             + PC1 + PC2
             + factor(Mask)
             + offset(log(population)),
             random = ~(1|state), family=negbin(1), data = mydata3)

summary(mod3$gam)
plot(mod3$gam, trans=exp, ylab="Mortality Rate Ratios", xlab="Weeks since the first death incidence", select=1)
```

(4) Weekly_death ~  s(days_since_first) + scaled independent variables + the 3 week lag positive test ratio + PC1 + PC2 + Mask indicator + random intercept

```{r,echo=FALSE}

mod4 = gamm4(weekly_death~ s(weeks_since_first, k=8, bs="cr") + factor(q_popdensity)
             + scale(Age_gp1) + scale(Age_gp3) + scale(Age_gp4) 
             + scale(poverty_perc) 
             + scale(Black.perc) + scale(Hispanic.or.Latino.perc)
             + scale(temp) + scale(perc)
             + scale(Bachelor_orhigher_perc) 
             + scale(bed_rate)  + scale(smoke_perc) + scale(obese_perc)  
             + ptest_weeks_ago3
             + PC1 + PC2
             + factor(Mask)
             + offset(log(population)),
             random = ~(1|state), family=negbin(1), data = mydata3)

summary(mod4$gam)
plot(mod4$gam, trans=exp, ylab="Mortality Rate Ratios", xlab="Weeks since the first death incidence", select=1)
```

