---
title: "Primary_Analysis_A"
author: "JLee"
date: "5/23/2020"
output:
  pdf_document: default
  word_document: default
  html_document: default
---

```{r,message=FALSE,include=FALSE,echo=FALSE}
library(mgcv);library(gamm4);library(ggplot2)
library(lubridate);library(dplyr);
library(boot)
library(table1)

# Data import
mydata <-  read.csv("/Users/jooyounglee/Dropbox/2020_covid19/covid19/Data/us_state_weekly_death_JHU.csv")[,-1]
mydata$date <- as.Date(mydata$date)
state_predic <-  read.csv("/Users/jooyounglee/Dropbox/2020_covid19/covid19/Data/us_state_predictors.csv")[,-1]

movement <- read.csv("/Users/jooyounglee/Dropbox/2020_covid19/covid19/Data/us_state_weekly_movement.csv")[,-1]
testing <- read.csv("/Users/jooyounglee/Dropbox/2020_covid19/covid19/Data/weekly_testing.csv")[,-1]
policy <-  read.csv("/Users/jooyounglee/Dropbox/2020_covid19/covid19/Data/state_weekly_policy.csv")[,-1]
movement$date <- as.Date(movement$date)
testing$date <- as.Date(testing$date)
policy$date <- as.Date(policy$date)

mydata <- merge(mydata, state_predic, by=c("state"))
mydata$date <- as.Date(mydata$date)
mydata <- mydata[order(mydata$state, as.Date(mydata$date)),]

```


```{r, include=FALSE, echo=FALSE}
# TABLE 1
#demographic_state[, 3:23] = sapply(demographic_state[, 3:23], as.numeric)

table1(~   Age_gp2 + Age_gp3 + Age_gp4 + 
         + density
         + temp + perc
         + Bachelor_orhigher_perc 
         + poverty_perc
         + Black.perc + Hispanic.or.Latino.perc
         + bed_rate + smoke_perc + obese_perc + diabetes_perc, data=state_predic)

```

# The death rate by state with the start of stay-at-home order and the end of it
```{r, message= FALSE, echo=FALSE}
mydata$death_rate <- mydata$weekly_death/mydata$population*100000

mydata0 <- merge(mydata, policy, by=c("state", "date"))

mydata0$date <- as.Date(mydata0$date)
mydata0$Start.Date <- as.Date(mydata0$Start.Date)
mydata0$End.Date <- as.Date(mydata0$End.Date)
qplot(date, death_rate, data=mydata0, geom="path", xlab="Date", ylab="Death rate") + guides(col = guide_legend(ncol=3)) + facet_wrap(.~state)+ geom_vline(aes(xintercept = Start.Date), data=mydata0, color="orange")+  geom_vline(aes(xintercept = End.Date), data=mydata0, color="purple") 
```

# The positive test rate by state with the start of stay-at-home order and the end of it
```{r, message= FALSE, echo=FALSE}
testing$rptest <- testing$pos_test/(testing$test)*100
mydata0 <- merge(mydata0, testing, by=c("state", "date"))

mydata0$date <- as.Date(mydata0$date)
mydata0$Start.Date <- as.Date(mydata0$Start.Date)
mydata0$End.Date <- as.Date(mydata0$End.Date)
qplot(date, rptest, data=mydata0, geom="path", xlab="Date", ylab="Positivce test rate rate") + guides(col = guide_legend(ncol=3)) + facet_wrap(.~state)+ geom_vline(aes(xintercept = Start.Date), data=mydata0, color="orange")+  geom_vline(aes(xintercept = End.Date), data=mydata0, color="purple") 
```


# The residential movement by state with the start of stay-at-home order and the end of it
```{r, message= FALSE, echo=FALSE}
mydata0 <- merge(mydata0, movement, by=c("state", "date"))
qplot(date, residential, data=mydata0, geom="path", xlab="Date", ylab="Residential Movement") + guides(col = guide_legend(ncol=3)) + facet_wrap(.~state)+ geom_vline(aes(xintercept = Start.Date), data=mydata0, color="orange")+  geom_vline(aes(xintercept = End.Date), data=mydata0, color="purple") 


```

```{r, include=FALSE, echo=FALSE}


# create lag for positive testing rate
lagk <- 3
testing$rptest <- testing$pos_test/(testing$test)*100

for (kk in (1:lagk)){
  create = paste0("ptest_weeks_ago",kk)
  for(i in (1:length(unique(testing$state)))){
    dates_before = testing[which(testing$state == unique(testing$state)[i]),]$date-7*kk
    ind = match(dates_before,testing[which(testing$state == unique(testing$state)[i]),]$date)
    testing[which(testing$state==unique(testing$state)[i]),create]=testing[which(testing$state==unique(testing$state)[i]),][ind,]$rptest
  }
}

# create lag for resident
lagk <- 3

for (kk in (1:lagk)){
  create = paste0("resident_weeks_ago",kk)
  for(i in (1:length(unique(movement$state)))){
    dates_before = movement[which(movement$state == unique(movement$state)[i]),]$date-7*kk
    ind = match(dates_before,movement[which(movement$state == unique(movement$state)[i]),]$date)
    movement[which(movement$state==unique(movement$state)[i]),create]=movement[which(movement$state==unique(movement$state)[i]),][ind,]$residential
  }
}

## Merging with time-varying variables - testing, movement, policy
mydata2 <- merge(mydata, testing, by=c("state", "date"))
mydata2 <- merge(mydata2, movement, by=c("state", "date"))

policy2 <- policy[, !names(policy) %in% c("Start.Date", "End.Date")]
policy2 <- policy2 %>% group_by(state) %>% mutate(shome_since_first = cumsum(stayathome))
policy2 <- policy2 %>% group_by(state) %>% mutate(mask_since_first = cumsum(Mask))
mydata2 <- merge(mydata2, policy2, by=c("state", "date"))
mydata2 <- as.data.frame(mydata2)

mydata2$rtest <- mydata2$test/(mydata2$population/1000)
#mydata$test_weeks_ago1 <- mydata$test_weeks_ago1/(mydata$population/1000)
#mydata$test_weeks_ago2 <- mydata$test_weeks_ago2/(mydata$population/1000)

mydata2 <- mydata2[mydata2$cum_death>=50, ]
mydata2 <- na.omit(mydata2)
mydata2 <- mydata2[mydata2$state!="Maine",]  ## Maine update the negative test results weekly or biweekly
# create # of weeks since first death (in our dataset)

mydata2$weeks_since_first = 0
for (i in 1:length(unique(mydata2$state))){
  for (k in 2:nrow(mydata2[which(mydata2$state==unique(mydata2$state)[i]),])){
  mydata2[which(mydata2$state==unique(mydata2$state)[i]),"weeks_since_first"][k]=
    (mydata2[which(mydata2$state==unique(mydata2$state)[i]),"date"][k]-mydata2[which(mydata2$state==unique(mydata2$state)[i]),"date"][1])/7
  }
}
```

# PCA 
```{r, echo=FALSE}
# create PCA for the movement data
pca_move <- prcomp(mydata2[,c("retail_rec", "grocery_pharmacy", "parks", "transit", "work", "residential")] # make sure col# correct
                   , center=TRUE, scale.=TRUE)
pca_move
summary(pca_move)
mydata3 <- cbind(mydata2, pca_move$x)

## truncate 
```

# Model 1

(1) Weekly_death ~ s(days_since_first) + scaled independent variables
```{r, echo=FALSE}

mod1 = gamm4(weekly_death~ s(weeks_since_first,k=8,bs='cr') + factor(q_popdensity)
             + scale(Age_gp1) + scale(Age_gp3) + scale(Age_gp4) 
             + scale(poverty_perc) #+ scale(log(susceptible_perc))
             + scale(Black.perc) + scale(Hispanic.or.Latino.perc)
             + scale(temp) + scale(perc)
             + scale(Bachelor_orhigher_perc) 
             + scale(bed_rate) + scale(smoke_perc) + scale(obese_perc)
             + offset(log(population)),
             random = ~(1|state), family=negbin(1), data = mydata3)

summary(mod1$gam)
plot(mod1$gam, trans=exp, ylab="Mortality Rate Ratios", xlab="Weeks since the first death incidence")
```

(2) Weekly_death ~ s(days_since_first) + scaled independent variables + positive_test_ratio with 3 weeks lag + random intercept 

```{r,echo=FALSE}

mod2 = gamm4(weekly_death~ s(weeks_since_first,k=8,bs='cr') + factor(q_popdensity)
             + scale(Age_gp1) + scale(Age_gp3) + scale(Age_gp4) 
             + scale(poverty_perc) 
             + scale(Black.perc) + scale(Hispanic.or.Latino.perc)
             + scale(temp) + scale(perc)
             + scale(Bachelor_orhigher_perc) 
             + scale(bed_rate) + scale(smoke_perc) + scale(obese_perc) 
             + rptest + ptest_weeks_ago1 + ptest_weeks_ago2 + ptest_weeks_ago3
             + offset(log(population)),
             random = ~(1|state), family=negbin(1), data = mydata3)

summary(mod2$gam)
plot(mod2$gam, trans=exp, ylab="Mortality Rate Ratios", xlab="Weeks since the first death incidence", select=1)
```

(3) Weekly_death ~  s(days_since_first) + scaled independent variables + the curent positive test ratio + the PCA of movement + the time since the mask policy with 3 levels + random intercept

```{r,echo=FALSE}

mydata3$f_mask <- mydata3$mask_since_first
mydata3$f_mask[mydata3$mask_since_first >=1 & mydata3$mask_since_first <=3 ] <- 1
mydata3$f_mask[mydata3$mask_since_first >=4 & mydata3$mask_since_first <=6 ] <- 2

mod3 = gamm4(weekly_death~ s(weeks_since_first, k=8, bs="cr") + factor(q_popdensity)
             + scale(Age_gp1) + scale(Age_gp3) + scale(Age_gp4) 
             + scale(poverty_perc) 
             + scale(Black.perc) + scale(Hispanic.or.Latino.perc)
             + scale(temp) + scale(perc)
             + scale(Bachelor_orhigher_perc) 
             + scale(bed_rate)  + scale(smoke_perc) + scale(obese_perc)  
             + rptest
             + PC1 + PC2 
             + factor(f_mask)
             + offset(log(population)),
             random = ~(1|state), family=negbin(1), data = mydata3)

summary(mod3$gam)
plot(mod3$gam, trans=exp, ylab="Mortality Rate Ratios", xlab="Weeks since the first death incidence", select=1)
```


(4) Weekly_death ~  s(days_since_first) + scaled independent variables + the positive test ratio with 1 week lag + the PCA of movement + the time since the mask policy with 3 levels + random intercept

```{r,echo=FALSE}

mod4 = gamm4(weekly_death~ s(weeks_since_first, k=8, bs="cr") + factor(q_popdensity)
             + scale(Age_gp1) + scale(Age_gp3) + scale(Age_gp4) 
             + scale(poverty_perc) 
             + scale(Black.perc) + scale(Hispanic.or.Latino.perc)
             + scale(temp) + scale(perc)
             + scale(Bachelor_orhigher_perc) 
             + scale(bed_rate)  + scale(smoke_perc) + scale(obese_perc)  
             + ptest_weeks_ago1
             + PC1 + PC2 
             + factor(f_mask)
             + offset(log(population)),
             random = ~(1|state), family=negbin(1), data = mydata3)

summary(mod4$gam)
plot(mod4$gam, trans=exp, ylab="Mortality Rate Ratios", xlab="Weeks since the first death incidence", select=1)
```

(5) Weekly_death ~  s(days_since_first) + scaled independent variables +  the positive test ratio with 2 week lag + the PCA of movement + the time since the mask policy with 3 levels + random intercept

```{r,echo=FALSE}

mod5 = gamm4(weekly_death~ s(weeks_since_first, k=8, bs="cr") + factor(q_popdensity)
             + scale(Age_gp1) + scale(Age_gp3) + scale(Age_gp4) 
             + scale(poverty_perc) 
             + scale(Black.perc) + scale(Hispanic.or.Latino.perc)
             + scale(temp) + scale(perc)
             + scale(Bachelor_orhigher_perc) 
             + scale(bed_rate)  + scale(smoke_perc) + scale(obese_perc)  
             + ptest_weeks_ago2 
             + PC1 + PC2 
             + factor(f_mask)
             + offset(log(population)),
             random = ~(1|state), family=negbin(1), data = mydata3)

summary(mod5$gam)
plot(mod5$gam, trans=exp, ylab="Mortality Rate Ratios", xlab="Weeks since the first death incidence", select=1)

```

(6) Weekly_death ~  s(days_since_first) + scaled independent variables +  the positive test ratio with 3 week lag + the PCA of movement + the time since the mask policy with 3 levels + random intercept

```{r,echo=FALSE}

mod5 = gamm4(weekly_death~ s(weeks_since_first, k=8, bs="cr") + factor(q_popdensity)
             + scale(Age_gp1) + scale(Age_gp3) + scale(Age_gp4) 
             + scale(poverty_perc) 
             + scale(Black.perc) + scale(Hispanic.or.Latino.perc)
             + scale(temp) + scale(perc)
             + scale(Bachelor_orhigher_perc) 
             + scale(bed_rate)  + scale(smoke_perc) + scale(obese_perc)  
             + ptest_weeks_ago3 
             + PC1 + PC2 
             + factor(f_mask)
             + offset(log(population)),
             random = ~(1|state), family=negbin(1), data = mydata3)

summary(mod5$gam)
plot(mod5$gam, trans=exp, ylab="Mortality Rate Ratios", xlab="Weeks since the first death incidence", select=1)

```


(7) Weekly_death ~  s(days_since_first) + scaled independent variables + positive test ratio with only 1-3 weeks lag without the current postive test ratio + the PCA of movement + the time since the mask policy with 3 levels + random intercept

```{r,echo=FALSE}

mod7 = gamm4(weekly_death~ s(weeks_since_first, k=8, bs="cr") + factor(q_popdensity)
             + scale(Age_gp1) + scale(Age_gp3) + scale(Age_gp4) 
             + scale(poverty_perc) 
             + scale(Black.perc) + scale(Hispanic.or.Latino.perc)
             + scale(temp) + scale(perc)
             + scale(Bachelor_orhigher_perc) 
             + scale(bed_rate)  + scale(smoke_perc) + scale(obese_perc)  
             + ptest_weeks_ago1 + ptest_weeks_ago2 + ptest_weeks_ago3 
             + PC1 + PC2 
             + factor(f_mask)
             + offset(log(population)),
             random = ~(1|state), family=negbin(1), data = mydata3)

summary(mod7$gam)
plot(mod7$gam, trans=exp, ylab="Mortality Rate Ratios", xlab="Weeks since the first death incidence", select=1)

```

(8) Weekly_death ~  s(days_since_first) + scaled independent variables + positive test ratio with only 1-3 weeks lag without the current postive test ratio + the residential movement with 3 weeks lag + the time since the mask policy with 3 levels  + random intercept

```{r,echo=FALSE}

mod5 = gamm4(weekly_death~ s(weeks_since_first, k=8, bs="cr") + factor(q_popdensity)
             + scale(Age_gp1) + scale(Age_gp3) + scale(Age_gp4) 
             + scale(poverty_perc) 
             + scale(Black.perc) + scale(Hispanic.or.Latino.perc)
             + scale(temp) + scale(perc)
             + scale(Bachelor_orhigher_perc) 
             + scale(bed_rate)  + scale(smoke_perc) + scale(obese_perc)  
             + ptest_weeks_ago1 + ptest_weeks_ago2 + ptest_weeks_ago3 
             + residential + resident_weeks_ago1 + resident_weeks_ago2 + resident_weeks_ago3 
             + factor(f_mask)
             + offset(log(population)),
             random = ~(1|state), family=negbin(1), data = mydata3)

summary(mod5$gam)
plot(mod5$gam, trans=exp, ylab="Mortality Rate Ratios", xlab="Weeks since the first death incidence", select=1)

```

(9) Weekly_death ~   linear days_since_first + scaled independent variables + positive test ratio with 3 weeks lag + random intercept + random slope for days_since_first

```{r,echo=FALSE}

mod9 = gamm4(weekly_death~ weeks_since_first + factor(q_popdensity)
             + scale(Age_gp1) + scale(Age_gp3) + scale(Age_gp4) 
             + scale(poverty_perc) 
             + scale(Black.perc) + scale(Hispanic.or.Latino.perc)
             + scale(temp) + scale(perc)
             + scale(Bachelor_orhigher_perc) 
             + scale(bed_rate)  + scale(smoke_perc) + scale(obese_perc)  
             + rptest + ptest_weeks_ago1 + ptest_weeks_ago2 + ptest_weeks_ago3 
             + offset(log(population)),
             random = ~(weeks_since_first|state), family=negbin(1), data = mydata3)

summary(mod9$gam)
```


(10) Weekly_death ~ linear days_since_first  + scaled independent variables + positive test ratio with only 1-3 weeks lag without the current postive test ratio +  random intercept + random slope for days_since_first

```{r,echo=FALSE}

mod10 = gamm4(weekly_death~ weeks_since_first  + factor(q_popdensity)
             + scale(Age_gp1) + scale(Age_gp3) + scale(Age_gp4) 
             + scale(poverty_perc) 
             + scale(Black.perc) + scale(Hispanic.or.Latino.perc)
             + scale(temp) + scale(perc)
             + scale(Bachelor_orhigher_perc) 
             + scale(bed_rate)  + scale(smoke_perc) + scale(obese_perc)  
             + ptest_weeks_ago1 + ptest_weeks_ago2 + ptest_weeks_ago3 
             + offset(log(population)),
             random = ~(weeks_since_first|state), family=negbin(1), data = mydata3)
summary(mod10$gam)
```


(11) Weekly_death ~ linear days_since_first  + scaled independent variables + positive test ratio with only 1-3 weeks lag without the current postive test ratio +  the PCA of movement + the time since the mask policy with 3 levels  +  random intercept + random slope for days_since_first

```{r,echo=FALSE}

mod11 = gamm4(weekly_death~ weeks_since_first  + factor(q_popdensity)
             + scale(Age_gp1) + scale(Age_gp3) + scale(Age_gp4) 
             + scale(poverty_perc) 
             + scale(Black.perc) + scale(Hispanic.or.Latino.perc)
             + scale(temp) + scale(perc)
             + scale(Bachelor_orhigher_perc) 
             + scale(bed_rate)  + scale(smoke_perc) + scale(obese_perc)  
             + ptest_weeks_ago1 + ptest_weeks_ago2 + ptest_weeks_ago3 
             + PC1 + PC2
             + factor(f_mask)
             + offset(log(population)),
             random = ~(weeks_since_first|state), family=negbin(1), data = mydata3)
summary(mod11$gam)
```

(12) Weekly_death ~ linear days_since_first  + scaled independent variables + positive test ratio with only 1-3 weeks lag without the current postive test ratio +  the residential movement with 3 weeks lag + the time since the mask policy with 3 levels + random intercept + random slope for days_since_first

```{r,echo=FALSE}

mod12 = gamm4(weekly_death~ weeks_since_first  + factor(q_popdensity)
             + scale(Age_gp1) + scale(Age_gp3) + scale(Age_gp4) 
             + scale(poverty_perc) 
             + scale(Black.perc) + scale(Hispanic.or.Latino.perc)
             + scale(temp) + scale(perc)
             + scale(Bachelor_orhigher_perc) 
             + scale(bed_rate)  + scale(smoke_perc) + scale(obese_perc)  
             + ptest_weeks_ago1 + ptest_weeks_ago2 + ptest_weeks_ago3 
             + residential + resident_weeks_ago1 + resident_weeks_ago2 + resident_weeks_ago3 
             + factor(f_mask)
             + offset(log(population)),
             random = ~(weeks_since_first|state), family=negbin(1), data = mydata3)
summary(mod12$gam)
```
